MIRSVM: Multi-Instance Support Vector Machine with Bag Representatives

### 2017-06-16 ArticleRevision/article/article.tex
Updated Gaussian formula + conclusion

One of the biggest advantages of the dual SVM formulation is the sparseness of the resulting model. This is because support vectors, instances that have their corresponding $\alpha_I \neq 0$, are only considered when forming the decision boundary. MIRSVM uses a Gaussian RBF kernel, given by Equation~\eqref{eq:gaussiankernel}, where $\sigma$ is the Gaussian shape parameter.
\begin{equation}
\label{eq:gaussiankernel}
\centering
\mathcal{K}(\bm x_i, \bm x_j) = e^{-\frac{\norm{x_i - x_j}^2}{2\sigma^2}}
\end{equation}
\begin{figure}[tpb!]
\centering
\small
\label{fig:mirsvm}
\begin{minipage}{\textwidth}
\adjustbox{scale=0.901}{
\begin{tikzpicture}[%
    node distance=4cm,
    on grid,
    auto
  ]
\node[block](A) {Initialize $\bm S$ randomly};
\node(B)[block, right of=A, text width=6em] {Train SVM using QP solver over $\bm S$, set $\bm S_{old} = \bm S$};
\draw[arrow] (A) -- (B);
\node(C)[block, right of=B,text width=6.3em] {Find optimal representatives, $\bm S$, for the current hyper-plane};
\draw[arrow] (B) -- (C);
\node(D) [decision,right of=C,fill=gray!55,text width=2cm]{while $\bm S$ has not changed, $\bm S \neq \bm S_{old}$};
\draw[arrow] (C) -- (D);
\node(E) [block,right of=D]{return $\bm \alpha,\, b, \bm S$};
\draw[arrow] (D) -- node {yes} (E);
\draw[arrow] (D.north) --+(0,0.1cm) -| (B.north) node[below,pos=0.25] {no} ;
\end{tikzpicture}}
\caption{MIRSVM flow diagram. This figure represents a summary of the steps performed by the MIRSVM algorithm. The representatives are first randomly initialized and continuously updated according to the current hyper-plane, which is found using a quadratic programming (QP) solver. Upon completion, the model is returned along with the optimal bag-representatives.}
\end{minipage}

@@

\section{Conclusion}\label{sec:conclusion}
This paper proposed a novel formulation and algorithm for the multiple-instance support vector machine problem, which optimizes bags classification via bag-representative selection. First, the primal formulation was posed and its dual was then derived and computed using a quadratic programming solver. This formulation was designed to utilize bag-level information and find an optimal separating hyperplane between bags, rather than individual instances, using the standard multi-instance assumption. The SMI assumption states that a bag is labeled positive if and only if at least one instance within a bag is positive, and is negative otherwise. The key features of the proposed algorithm MIRSVM are its ability to identify instances within positive and negative bags, i.e. the support vectors or representatives, that highly impact the decision boundaries as well as avoiding uncertainties and issues caused by techniques that flatten, subset, or under-represent positive instances within positively labeled bags. Additionally, it exhibits desirable scalability, making it suitable for large-scale learning tasks.

The experimental study showed the better performance of MIRSVM compared with multi-instance support vector machines, traditional multi-instance learners, as well as with ensemble methods. The results, according to a variety of performance metrics, were compared and further validated using statistical analysis with non-parametric tests which highlight the advantages of using bag-level based and ensemble learners such as Bagging and Stacking, while showing the instance-level based learners performed poorly in comparison or were deemed as strongly biased and unstable classifiers. On the contrary, our proposal MIRSVM performs statistically better, neither compromising accuracy nor run-time while displaying a robust performance across all of the evaluated datasets.

>>>>

One of the biggest advantages of the dual SVM formulation is the sparseness of the resulting model. This is because support vectors, instances that have their corresponding $\alpha_I \neq 0$, are only considered when forming the decision boundary. MIRSVM uses a Gaussian RBF kernel, given by Equation~\eqref{eq:gaussiankernel}, where $\sigma$ is the Gaussian shape parameter.
\begin{equation}
\label{eq:gaussiankernel}
\centering
\mathcal{K}(\bm x_i, \bm x_j) = e^{-\frac{\norm{\bm x_i - \bm x_j}^2}{2\sigma^2}}
\end{equation}
\begin{figure}[tpb!]
\centering
\small
\label{fig:mirsvm}
\begin{minipage}{\textwidth}
\adjustbox{scale=0.901}{
\begin{tikzpicture}[%
    node distance=4cm,
    on grid,
    auto
  ]
\node[block](A) {Initialize $\bm S$ randomly};
\node(B)[block, right of=A, text width=6em] {Train SVM using QP solver over $\bm S$, set $\bm S_{old} = \bm S$};
\draw[arrow] (A) -- (B);
\node(C)[block, right of=B,text width=6.3em] {Find optimal representatives, $\bm S$, for the current hyper-plane};
\draw[arrow] (B) -- (C);
\node(D) [decision,right of=C,fill=gray!55,text width=2cm]{while $\bm S$ has not changed, $\bm S \neq \bm S_{old}$};
\draw[arrow] (C) -- (D);
\node(E) [block,right of=D]{return $\bm \alpha,\, b, \bm S$};
\draw[arrow] (D) -- node {yes} (E);
\draw[arrow] (D.north) --+(0,0.1cm) -| (B.north) node[below,pos=0.25] {no} ;
\end{tikzpicture}}
\caption{MIRSVM flow diagram. This figure represents a summary of the steps performed by the MIRSVM algorithm. The representatives are first randomly initialized and continuously updated according to the current hyper-plane, which is found using a quadratic programming (QP) solver. Upon completion, the model is returned along with the optimal bag-representatives.}
\end{minipage}

@@

\section{Conclusion}\label{sec:conclusion}
This paper proposed a novel formulation and algorithm for the multiple-instance support vector machine problem, which optimizes bag classification via bag-representative selection. First, the primal formulation was posed and its dual was then derived and solution computed using a quadratic programming solver. This formulation was designed to utilize bag-level information and find an optimal separating hyperplane between bags, rather than individual instances, using the standard multi-instance assumption. The SMI assumption states that a bag is labeled positive if and only if at least one instance within a bag is positive, and is negative otherwise. The key features of the proposed algorithm MIRSVM are its ability to identify instances within positive and negative bags, i.e. the support vectors or representatives, that highly impact the decision boundary and margin, as well as avoiding uncertainties and issues caused by techniques that flatten, subset, or under-represent positive instances within positively labeled bags. Additionally, it exhibits desirable convergence and scalability, making it suitable for large-scale learning tasks.

The experimental study showed the better performance of MIRSVM compared with state-of-the-art multi-instance support vector machines, traditional multi-instance learners, as well as ensemble methods. The results, according to a variety of performance metrics, were compared and further validated using statistical analysis with non-parametric tests which highlight the advantages of using bag-level based and ensemble learners, such as Bagging and Stacking, while showing the instance-level based learners performed poorly in comparison or were deemed as strongly biased and unstable classifiers. On the contrary, our proposal MIRSVM performs statistically better, neither compromising accuracy nor run-time while displaying a robust performance across all of the evaluated datasets.
